Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Accepting Sou-Cheng's and Aleksei's comments.

We draw on our collective experiences as
members of the academic QMC community, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists. Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.   \FJHNote{Sou-Cheng, not sure what this means.} \SCComment{Fred, more elaboration here, but maybe it fits better in Section 3: In particular, QMCPy is not just a QMC software containing our own algorithms; from the outset, it is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times. Using object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects essential in multiple integration (see Sect.\ \ref{CDHJS_sec:integrated}). Developers of other (Q)MC software could easily extend these objects to incorporate their existing or working algorithms by a combination of mechanism: object inheritance; Python package inclusion; direct code translation; interfaces to compiled C or Java libraries. Hence QMCPy has been able to provide access to an increasing number of well-established (Q)MC algorithmic or application software.}

@@

@@

Great QMC software eliminates the need for researchers to resort to  ``do-it-yourself'' for established algorithms.  The less code we have to write, the fewer errors. \AGSComment{Aleksei: Put in previous subsection.}

@@

Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.  Stephen Joe and Frances Kuo \cite{JoeKuo03,JoeKuo08,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.  Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria. During a lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.

Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.  However, most of these implementations in large libraries could provide more flexible \SCComment{what does more flexible mean?} LD offerings.  Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings. \SCComment{Other established software such as SAS expand offerings of trustworthy and efficient analytics through continuous internal implementation as well as flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}. Such a latter software development strategy presents opportunities for}  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to demonstrate how their libraries can connect well with other software environments.  An example of QMCPy connecting with other software is given in the next subsection.

@@

To put $\mu$ into the form that is directly accessible to QMC methods requires a variable transformation, $\bsPsi$, such that
\begin{equation*}
    \bsPsi((0,1)^d) = \calT, \qquad f(\bsx) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial\bsx}}.
\end{equation*}
Such a transformation is typically non-unique, and the choice of a good one is equivalent to importance sampling.  If $\lambda$ is a PDF for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.  In this case, one might choose $\bsPsi$ such that $\bsT \sim \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$. \AGSComment{Aleksei: Software such as SciPy and QMCPy facilitate such variable transformations by providing default $\bsPsi$ for common $\bsT$. This alleviates the burden on the user to determine a transformation and implement the Jacobian. }

@@

\SCComment{The above examples and discussion also underscore the importance of continuous testing in QMC software development. This involves creating a comprehensive test suite that covers a wide range of scenarios and input parameters. Additionally, QMC software should be benchmarked against known results to ensure that it is both accurate and efficient.}

@@

Efficiency gains in QMC software can also result from clever algorithm choices. \SCComment{Adaptive QMC algorithms as in \cite{JimHic16a, HicJim16a, HicEtal17a} can be used to dynamically increase the number of sampling points based on estimated error bounds. They can lead to significant efficiency gains by reducing the number of unnecessary function evaluations.} The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.  This reduces the computations from what would be $\calO(n^3)$ for arbitrary covariance kernels to only $\calO(n \log n)$ for these well-chosen kernels. Consequently, computing the stopping criterion is roughly the same order of magnitude as computing the solution.

@@

    \item \SCComment{Let's publish our QMC  software with journals such as ACM TOMS, JORS, etc. to obtain code review and
    feedback.}

    \item \SCComment{When we find QMC scientific software useful in our research or applications, let's encourage and recognize such software by giving them proper citations~\cite{smith2016software}. }

    \item \SCComment{Encourage QMC community members to learn about software development through interdisciplinary education, internship, or hackathon. }

@@

Encouraging students and postdocs to invest time in developing QMC software will require us as their mentors valuing software as a research output.  Great software should be published with a digital object identifier (DOI).  \SCComment{The previous sentence is about publishing software. The next point is about a career path.}  The term research software engineer (RSE) is becoming a more valued vocation.  According to the website for  Research Software Engineers International, \url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.}

@@

\SCComment{Change `sobol' to `Sobol' in reference items 29 and 30. Change `scipy' to `SciPy' in 44 and 45. }
\bibliographystyle{spmpsci}
\bibliography{FJH23,FJHown23,sc}

>>>>

We draw on our collective experiences as
members of the academic QMC community, developers of open-source and commercial scientific software, and users of a wide range of scientific software as research scientists or data scientists. Many of the insights that we have gained have come through our development of the Guaranteed Automatic Integration Library (GAIL)~\cite{ChoEtal22a} in MATLAB and the Quasi-Monte Carlo Python library (QMCPy)~ \cite{QMCPy2020a}.

@@

Great QMC software eliminates the need for researchers to resort to  ``do-it-yourself'' for established algorithms.  The less code we have to write, the fewer errors.

@@

@@

Lattice and digital sequences are not unique, and improved or alternate versions continue to be constructed.  Stephen Joe and Frances Kuo \cite{JoeKuo03,JoeKuo08a,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of early LD sequence software.  Pierre L'Ecuyer and collaborators have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user-defined criteria. During a lunch at MCQMC 2022, it was proposed that we  agree on a consistent format for storing the parameters that define good LD sequences so that they can be shared across software libraries.

Users computing solutions in comprehensive software environments, such as NAG, MATLAB, R, SciPy, and TensorFlow, have convenient access to LD sequence generators.  However, most of these implementations in large libraries would benefit a wider menu of LD offerings, including custom generators from LatNetBuilder or other sources.  Some large environments such as  Dakota \cite{DakotaUsersManual} have quite limited LD generators and should expand their offerings. SAS facilitates flexible integration of third-party open-source software; see \url{https://www.sas.com/en_us/software/viya/open.html}. This presents opportunities for  developers of smaller libraries with more QMC features, like BRODA and QMCPy, to connect their libraries with larger software environments. An example of QMCPy connecting with other software is given in the next subsection.

In fact, QMCPy is designed not simply to contain our own algorithms.  It is designed to serve as a middleware framework, allowing users to tap into a collection of QMC libraries originally written in various languages by different researchers at different times. Using object-oriented design, QMCPy specifies a few  extensible high-level mathematical objects as described in this section and in \cite{ChoEtal22a}. Developers of other (Q)MC software can  extend these objects to incorporate their existing algorithms by a combination of mechanisms: object inheritance, Python package inclusion, direct code translation, and/or interfaces to compiled C or Java libraries.

@@

To put $\mu$ into the form that is directly accessible to QMC methods requires a variable transformation, $\bsPsi$, such that
\begin{equation*}
    \bsPsi((0,1)^d) = \calT, \qquad f(\bsx) = g(\bsPsi(\bsx)) \lambda(\bsPsi(\bsx)) \abs{\frac{\partial \bsPsi}{\partial\bsx}}.
\end{equation*}
Such a transformation is typically non-unique, and the choice of a good one is equivalent to importance sampling.  If $\lambda$ is a PDF for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.  In this case, one might choose $\bsPsi$ such that $\bsT \sim \bsPsi(\bsX)$, in which case $\lambda(\bsPsi(\bsx)) \abs{\partial \bsPsi/\partial\bsx} = 1$ and $f(\bsx) = g(\bsPsi(\bsx))$. Software such as SciPy and QMCPy facilitate such variable transformations by providing default $\bsPsi$ for common $\bsT$. This alleviates the burden on the user to determine a transformation and implement the Jacobian.

@@

The above examples and discussion also underscore the importance of continuous testing in QMC software development. This involves creating a comprehensive test suite that covers a wide range of scenarios and input parameters. Additionally, QMC software should be benchmarked against known results to ensure that it is both accurate and efficient.

@@

Efficiency gains in QMC software can also result from clever algorithm choices. Adaptive QMC algorithms as in \cite{JimHic16a, HicJim16a, HicEtal17a, RatHic19a,JagHic22a} can be used to dynamically increase the number of sampling points based on estimated error bounds. They can lead to significant efficiency gains by reducing the number of unnecessary function evaluations. The Bayesian stopping rules for lattice and Sobol sequences in \cite{RatHic19a,JagHic22a} choose covariance kernels that match the corresponding LD sequence.  This reduces the computations from what would be $\calO(n^3)$ for arbitrary covariance kernels to only $\calO(n \log n)$ for these well-chosen kernels. Consequently, computing the stopping criterion is roughly the same order of magnitude as computing the solution.

@@

    \item Let's publish our QMC  software with journals such as ACM TOMS, JORS, etc. to obtain code review and
    feedback.

    \item When we find QMC scientific software useful in our research or applications, let's encourage and recognize such software by properly citing it~\cite{smith2016software}.

    \item Encourage QMC community members to learn about software development through interdisciplinary education, internships, or hackathons.

@@

Encouraging students and postdocs to invest time in developing QMC software will require us as their mentors valuing software as a research output.  Great software should be published with a digital object identifier (DOI).

The term research software engineer (RSE) is becoming a more valued vocation.  According to the website for  Research Software Engineers International, \url{https://researchsoftware.org}, \emph{Research Software Engineers are people who combine professional software expertise with an understanding of research.}

@@

\bibliographystyle{spmpsci}
\bibliography{FJH23,FJHown23,sc}
